Masterclass Assignment: Real-time Global Illumination

### 2018-01-13 docs/paper.tex
spell checked some of it

The first step of implementing VXGI was to create a voxelized representation of the albedo and surface normal information in the scene. The method used in the source material \cite{crassinneyretsainzgreeneisemann2011} is explained in great detail in an OpenGL Insights chapter titled \textit{Octree-Based Sparse Voxelization Using the GPU Hardware Rasterizer} \cite{crassingreen2012}, but the basic concept is as followed:
\begin{enumerate}
	\item Generate three projection matrices for orthogonal projections, covering the scene geometry equally, and aligned to each of the coordinate axes.
	\item Draw the scene with depth testing disabled, ensuring fragments are generated for every triangle.
	\item For each triangle, determine which of the coordinate axes is most aligned with the surface normal, thus generating the maximum number of fragments when transformed using the corresponding projection matrix for that axis.
	\item Using either the NVIDIA OpenGL extension \textit{GL CONSERVATIVE RASTERIZATION NV}, or a geometry shader, enlarge each triangle so that fragments are generated regardless of whether it covers a pixels center. This helps to reduce 'cracks' in the voxelisation on surfaces that are angled away from the coordinate axes.
	\item Using the fragment coordinates, depth, and the axis chosen in step 3, infer the texel coordinate that the fragment corresponds to in the target 3D textures. Store the fragment albedo and normal information in 3D textures using a moving average.
\end{enumerate}
Whilst this technique is somewhat easy to explain, I had not previously used the required Image Load/Store commands so I referred to an implementation I found on GitHub \cite{voxelization} to guide me on the OpenGL side of things.

The next step was to inject the scene's emissive values into a 3D texture. This is achieved by raytracing towards the light, for each of the non-transparent texels in the voxelised textures. This is where my implementation differs from the source material \cite{crassinneyretsainzgreeneisemann2011} as I opted to use a dense 3D texture as opposed to a sparse voxel octree. Whilst this was significantly easier to set up, it also severly impacts the speed at which raytracing can be performed through this 3D texture (See Figure~\ref{figure:performance} for performance metrics). The source material \cite{crassinneyretsainzgreeneisemann2011} also details a method for mip-mapping this texture anisotropically along each of the coordinate axes, however I opted to use OpenGL's \lstinline{genMipMaps} method to save time.

@@

To calculate the the direct lighting contribution I referred to tutorial materian from LearnOpenGL \cite{learnPBR} to ensure that the PBR textures passed through the G-Buffer are combined in a physically plausible way. I referred again to VCTRenderer \cite{Villegas2016} to implement soft shadows into the direct lighting component, which involves calculating a fragment's occlusion by tracing a cone back towards the light source and accumulating the opacity values at each step.

@@

Figure~\ref{figure:fullrender} shows a series of screenshots of my final results with all of the lighting passes enabled. Whilst I am pleased with the outcome, there are also a number of visual artefacts that I discuss in Section~\ref{section:improvements}. Figure~\ref{figure:performance} shows some performance metrics for the individual lighting passes at different voxel resolutions. Note that memory usage is high, even at low resolutions, as I opted to use dense 3D textures as opposed to sparse 3D textures. The time taken to perform the light injection pass also scales quite poorly at high resolutions because of this. Despite this, moving around the scene maintains a smooth 60 frames per second (limited by the OpenGL context), providing the user does not move the light. The voxelization and shading passes seem to scale very well with increased resolution, taking on average 2 and 4.5 miliseconds respectively at all of the tested resolutions.

@@

There are also a number of visual artefacts, the most noticable of which being the cracks in the voxelization as seen in Figure~\ref{figure:cracks}. I understand this to be due to incorrect conservative rasterization, however I have the NVIDIA \textit{GL CONSERVATIVE RASTERIZATION NV} extension enabled and have confirmed it to be working, so I believe it could be an issue with using the extension whilst having multisampling enabled (as it is currently). I found that using a geometry shader to achieve the same result led to quite unusable artefacting.

Another artefact is quite visible noise in the indirect lighting component, as seen in Figure~\ref{figure:noise}. I am not sure what causes this, though I suspect it could be because the hemisphere of indirect cones are aligned with the normal-mapped surface normal as opposed to the surface normal recieved from the geometry, which results in excessive variance between indirect lighting calculations for neighboring fragments. Banding artefacts are visible in the specular component, as seen in Figure~\ref{figure:banding}, which occurs at certain viewing angles and at certain values for the specular aperture multiplier. I suspect this is a combination of excessive variance in surface normals, as mentioned previously, but could also be a result of limited numerical precision in calculating the specular component.

>>>>

The first step of implementing VXGI was to create a voxelized representation of the albedo and surface normal information in the scene. The method used in the source material \cite{crassinneyretsainzgreeneisemann2011} is explained in great detail in an OpenGL Insights chapter titled \textit{Octree-Based Sparse Voxelization Using the GPU Hardware Rasterizer} \cite{crassingreen2012}, but the basic concept is as followed:
\begin{enumerate}
	\item Generate three projection matrices for orthogonal projections, covering the scene geometry equally, and aligned to each of the coordinate axes.
	\item Draw the scene with depth testing disabled, ensuring fragments are generated for every triangle.
	\item For each triangle, determine which of the coordinate axes is most aligned with the surface normal, thus generating the maximum number of fragments when transformed using the corresponding projection matrix for that axis.
	\item Using either the NVIDIA OpenGL extension \textit{GL CONSERVATIVE RASTERIZATION NV}, or a geometry shader, enlarge each triangle so that fragments are generated regardless of whether it covers a pixels center. This helps to reduce 'cracks' in the voxelization on surfaces that are angled away from the coordinate axes.
	\item Using the fragment coordinates, depth, and the axis chosen in step 3, infer the texel coordinate that the fragment corresponds to in the target 3D textures. Store the fragment albedo and normal information in 3D textures using a moving average.
\end{enumerate}
Whilst this technique is somewhat easy to explain, I had not previously used the required Image Load/Store commands so I referred to an implementation I found on GitHub \cite{voxelization} to guide me on the OpenGL side of things.

The next step was to inject the scene's emissive values into a 3D texture. This is achieved by raytracing towards the light, for each of the non-transparent texels in the voxelized textures. This is where my implementation differs from the source material \cite{crassinneyretsainzgreeneisemann2011} as I opted to use a dense 3D texture as opposed to a sparse voxel octree. Whilst this was significantly easier to set up, it also severely impacts the speed at which raytracing can be performed through this 3D texture (See Figure~\ref{figure:performance} for performance metrics). The source material \cite{crassinneyretsainzgreeneisemann2011} also details a method for mip-mapping this texture anisotropically along each of the coordinate axes, however I opted to use OpenGL's \lstinline{genMipMaps} method to save time.

@@

To calculate the the direct lighting contribution I referred to tutorial material from LearnOpenGL \cite{learnPBR} to ensure that the PBR textures passed through the G-Buffer are combined in a physically plausible way. I referred again to VCTRenderer \cite{Villegas2016} to implement soft shadows into the direct lighting component, which involves calculating a fragment's occlusion by tracing a cone back towards the light source and accumulating the opacity values at each step.

@@

Figure~\ref{figure:fullrender} shows a series of screenshots of my final results with all of the lighting passes enabled. Whilst I am pleased with the outcome, there are also a number of visual artefacts that I discuss in Section~\ref{section:improvements}. Figure~\ref{figure:performance} shows some performance metrics for the individual lighting passes at different voxel resolutions. Note that memory usage is high, even at low resolutions, as I opted to use dense 3D textures as opposed to sparse 3D textures. The time taken to perform the light injection pass also scales quite poorly at high resolutions because of this. Despite this, moving around the scene maintains a smooth 60 frames per second (limited by the OpenGL context), providing the user does not move the light. The voxelization and shading passes seem to scale very well with increased resolution, taking on average 2 and 4.5 milliseconds respectively at all of the tested resolutions.

@@

There are also a number of visual artefacts, the most noticeable of which being the cracks in the voxelization as seen in Figure~\ref{figure:cracks}. I understand this to be due to incorrect conservative rasterization, however I have the NVIDIA \textit{GL CONSERVATIVE RASTERIZATION NV} extension enabled and have confirmed it to be working, so I believe it could be an issue with using the extension whilst having multisampling enabled (as it is currently). I found that using a geometry shader to achieve the same result led to quite unusable artefacting.

Another artefact is quite visible noise in the indirect lighting component, as seen in Figure~\ref{figure:noise}. I am not sure what causes this, though I suspect it could be because the hemisphere of indirect cones are aligned with the normal-mapped surface normal as opposed to the surface normal recieved from the geometry, which results in excessive variance between indirect lighting calculations for neighbouring fragments. Banding artefacts are visible in the specular component, as seen in Figure~\ref{figure:banding}, which occurs at certain viewing angles and at certain values for the specular aperture multiplier. I suspect this is a combination of excessive variance in surface normals, as mentioned previously, but could also be a result of limited numerical precision in calculating the specular component.
